%\newpage
\section*{Quizzes}

Throughout the semester, we will occasionally have a quiz in class or
recitation --- \emph{don't be afraid of quizzes: they are a great way
  to get participation points!}  Let's make sure you are all set.

\begin{part}\TAGS{course-policies}
  Point your browser to the
  \href{http://cs.cmu.edu/~15122/quiz.shtml}{quiz page}
  (\url{http://cs.cmu.edu/~15122/quiz.shtml}) and click on the
  \fbox{\textbf{Test}} button.  Several things could happen:
  \begin{enumerate}
  \item%
    You get an error message.  Go back to the
    \href{http://cs.cmu.edu/~15122/quiz.shtml}{quiz page} and follow
    the instructions under ``Common misconfigurations''.  Then try again.
  \item%
    You end up in your personal Google account.  Click on the top
    right icon.  If your andrew email address is listed, click on it
    and continue with step (c).  If you don't see your andrew email
    address, click on ``Add account'' to enter it and then continue
    with step (c).
  \item%
    You get to a Google sign-in page.  Authenticate with your
    \textbf{andrew email address}.  Then proceed to step (d).
  \item%
    You are redirected to CMU's \href{https://login.cmu.edu/}{web
      login page}.  Authenticate and proceed to step (e).
  \item%
    You see a Google form entitled ``Testing your configuration'' ---
    good!  Click on the radio button and then \fbox{\textbf{submit}}.
    You are ready to take quizzes!
  \end{enumerate}
  If you get stuck at any point, ask a TA to help you.
\end{part}

\bigskip
At this point you are done with the lab, and you are free to go.
