\part[1]\TAGS{ints}
For each of the following statements, determine whether the statement
is true or false in C0.  If it is true, explain why in one
sentence.  If it is false, give a counterexample to illustrate why the
statement is false.

\bigskip
For every \lstinline'int x, y': \hspace{0.25in}
          if \lstinline'x < y', then \lstinline'x + 1 <= y'.
\begin{framed}
\ifprintanswers{\color{\answerColor}
  True since, $x < y$ always implies $x+1 \le y$
}\else~\vspace{1in}\fi
\end{framed}

\RUBRIC
(x < y implies x+1 <= y)
TAGS: ints

+0.1pt True
+0.15pt --EITHER-- Fully correct explanation including handling edge case (mentions overflow and/or y == int_max())
+0.05pt --OR-- Partialy correct explanation based on integer arithmetic without considering edge cases
ENDRUBRIC

\medskip
For every \lstinline'int x': \hspace{0.25in}
          \lstinline'x >> 1' is equivalent to \lstinline'x / 2'.
\begin{framed}
\ifprintanswers{\color{\answerColor}
  False.  Counterexample: any odd negative number.
}\else~\vspace{1in}\fi
\end{framed}

\medskip
For every \lstinline'int x, y, z': \hspace{0.25in}
          \lstinline'(x + y) * z' is equivalent to \lstinline'z * y + x * z'.
\begin{framed}
\ifprintanswers{\color{\answerColor}
  True, since distributivity and commutativity hold in modular arithmetic.
}\else~\vspace{1in}\fi
\end{framed}

\medskip
For every \lstinline'int x, y': \hspace{0.25in}
 \lstinline'x < y' is equivalent to \lstinline'x - y < 0'.
\begin{framed}
\ifprintanswers{\color{\answerColor}
  False.  Counterexample: \lstinline'x = int_min()', %
  \lstinline'y = int_max()' since \lstinline'x - y == 1'.
}\else~\vspace{1in}\fi
\end{framed}

\RUBRIC
Part (b)
TAGS: ints

Gradescope rubric:
+ 0.25 pts Box 1: (now separate rubric item)
+ 0.25 pts Box 2: false + counterexample
+ 0.25 pts Box 3: true + appeal to axioms that hold true for modular arithmetic (distributivity, commutativity)
+ 0.25 pts Box 4: false + counterexample

Commentary:
1/4 point each. Student must get true/false right
AND the explanation or counter-example. If students says true
correctly but gives wrong reason, it's wrong.

  - (now separate rubric item)
  - False, any ODD negative number
  - True, principle
  - False, lots of examples, x = int_min(), y = int_max() is one
    example (x - y = 1)
ENDRUBRIC
