\section*{Making C0 work the next time you do something}
\enlargethispage{3em}

The steps executed in Task~\textbf{(3.a)} only set the path variable
for \lstinline[language={[coin]C}]'cc0' and
\lstinline[language={[coin]C}]'coin' for the current session, and you
don't want to have to repeat that every time you log in.  To set the
path variable permanently, follow these instructions:


\vspace*{-1ex}
\begin{part}
  Go to \url{http://c0.typesafety.net/tutorial/C0-at-CMU.html} and do
  the \textbf{second} step, \emph{Setting up your environment}.
\end{part}

\vspace*{-1ex}
\begin{part}
  Test that this works by \textbf{closing and reopening your
    terminal}, and typing

  \lstinline[language={[coin]C}]'% coin -d'

  If the command above correctly starts the interpreter, then you're
  set up!  \textbf{Don't forget to log out again before you leave.}
\end{part}

\vspace*{-1ex}
\begin{part}
  \newcommand{\lapsessionTime}{Wednesday from 4:30 to 6:30}
  \newcommand{\lapsessionPlace}{DH 2315}

  If you have your laptop, you can try to work through the first step
  of C0 at CMU, \emph{Connecting through SSH}, on your
  laptop. However, we'll have a laptop session on \lapsessionTime{} in
  \lapsessionPlace{} for you to get your laptop set up.
\end{part}

\begin{comment}
Exercise: %

As you type code in C0, you will likely get a lot of syntax errors, so
you need to read them carefully to figure out what the compiler is
telling you. In this final exercise, use the text editor you chose to
edit the factorial.c0 file and change one instruction to introduce a
syntax error. Save and exit from the editor and then use the compiler
or interpreter to generate a syntax error. Read the error and see if
it makes sense based on the edit you did. Does the error include a
line number? If so, was the error on that line number, or before that
line number? Go back into the editor and fix the program and make sure
it compiles and runs successfully.  FUN ACTIVITY If someone sitting
next to you is also done, edit your factorial function so it generates
a syntax error, and see if they can figure out what you did!  4 Handin
There is no hand in required for Lab 0. Your teaching assistant will
give you instructions on what to do next.
\end{comment}
