\documentclass[12pt]{exam}
\newcommand{\hwnumber}{1}
\newcommand{\duedate}{\formatdate{2}{9}{\YEAR}} % day-month-year
\newcommand{\shortdesc}{This written homework is the first of two
  homeworks that will introduce you to the way we reason about C0 code
  in 15-122.  It also makes sure that you have a good understanding of
  key course policies.}

\input{inc/preamble}
\begin{document}
\input{inc/frontmatter}

\begin{questions}
\inputQuestion[questions/oli/]{pretest}
\inputQuestion{running-c0}  % See file
\inputQuestion{policies} % See file
\inputQuestion[questions/pre-post/]{main1}  % F19 F18 F17 F16 S16
%\inputQuestion[questions/pre-post/]{main2}  % S19 S18 S17
\inputQuestion[questions/loops/]{main} % See file
%\inputQuestion{loop-invariant}
\inputQuestion[questions/function-correct/]{main1}  % F19 F18 F17 F16 S16
                                                    % and see file
%\inputQuestion[questions/function-correct/]{main2}  % S19 S18 S17                                                    % and see file
%\inputQuestion[questions/loop-termination/]{main1}
%\inputQuestion[questions/loop-termination/]{main2}
\end{questions}

\RUBRIC

Learning Objectives (AKA Evaluation Summary)
* course policies
* contracts
* tracing code
* proving correctness
* logical vs. operational reasoning
* termination

Gradescope rubric (no points, first two are always the same):
-0pt EXCELLENT JOB!!!
-0pt There are several concepts you have doubts about (see below): please consider going to office hours.

-0pt You will need to review the course policies: you don't want to get yourself into trouble!
-0pt It looks like you don't have a good grasp of function contracts.
-0pt You seem to have difficulties tracing the value of variables through a loop.
-0pt You might not fully understand how to prove a function correct.
-0pt You seem a bit confused about logical versus operational reasoning. See the post on Diderot called "Tips on Reasoning about Code".
-0pt You seem a bit confused about termination.

ENDRUBRIC

\inputQuestion{inc/backmatter}
\end{document}
